---
title: "Hub genes and cytokines influences on MRSA"
date: "5/20/2022"
author: Yenlin Lai
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read data

```{r}
# importing required packages
library(readxl)

#get your working directionary
getwd()

#change to your file path
MRSA_status  <- read_excel("./MRSA_Decrypted.xlsx", 1)
B_cell <- read_excel("./MRSA_Decrypted.xlsx", 2)
T_cell  <- read_excel("./MRSA_Decrypted.xlsx", 3)
Cytokine  <- read_excel("./MRSA_Decrypted.xlsx", 4)
```

## Combining the data with all rows

```{r}
library(dplyr)

df <- MRSA_status %>% 
  left_join(B_cell, by = c("SID" = "SID")) %>% 
  left_join(T_cell, by = c("SID" = "SID"), suffix = c("_B_Cell", "_T_Cell")) %>% 
  left_join(Cytokine, by = c("SID" = "SID"))

head(df)
```

1. What hub genes and cytokines are significantly different between persistent and resolving outcome of MRSA infection in the cohort of 58 subjects? 
a. Do these hub genes and cytokines influence outcome of MRSA individually?
b. Do these hub genes and cytokines influence outcome of MRSA together? 

```{r}
library(dplyr)
df1 <- 
  df %>% select(-SID)

###First Method: Lasso###
##individually##
library(glmnet)
x <- df1 %>% select(-Status) %>% as.matrix()
y <- df1 %>% select(Status) %>% unlist() %>% as.numeric() 
fit_lasso <- 
  cv.glmnet(x, y, family = "binomial", type.measure = "deviance", alpha = 1)
plot(fit_lasso)
fit_lasso$lambda.min
fit_lasso$lambda.1se
cof <- coef(fit_lasso$glmnet.fit, s=fit_lasso$lambda.min)
sum(cof != 0)
as.matrix(cof)[as.matrix(cof) != 0,]
#We found 6 significant variables: Hub13_B_Cell, Hub2_T_Cell, Hub18_T_Cell, CYTO1, CYTO2, CYTO3.#

##Fit these variables to logistic regression model##
logistic_lasso <- glm(Status ~ (Hub13_B_Cell + Hub2_T_Cell + Hub18_T_Cell + 
                            CYTO1 + CYTO2 + CYTO3), data = df1,
                family = "binomial")
summary(logistic_lasso)
step_ll <- step(logistic_lasso)
summary(step_ll)
pchisq(step_ll$deviance, step_ll$df.residual, lower.tail = FALSE)
#This stepwise model is significant#
#We found 5 significant variables after using stepwise: Hub13_B_Cell, Hub18_T_Cell, CYTO1, CYTO2, CYTO3. These are the variables which influence outcome of MRSA individually the most.#

##together##
logistic_lasso_tog <- glm(Status ~ ((Hub13_B_Cell + Hub2_T_Cell +
                                       Hub18_T_Cell)*(CYTO1 + CYTO2 + CYTO3)),
                          data = df1, family = "binomial")
summary(logistic_lasso_tog)
anova(logistic_lasso, logistic_lasso_tog, test = "Chi")
##The interaction terms are not significant. These hub genes and cytokines does not influence outcome of MRSA together.##

###Second Method: Random Forest###
library(randomForest)
set.seed(402)
fmod <- randomForest(Status ~., data = df1)
importance(fmod)
#Set the threshold to be 0.4, we found 11 significant variables: Hub13_B_Cell, Hub2_T_Cell, Hub4_T_Cell, Hub10_T_Cell, Hub11_T_Cell, Hub13_T_Cell, Hub16_T_Cell, Hub18_T_Cell, CYTO1, CYTO2, CYTO3.#

##Fit these variables to logistic regression model##
logistic_rf <- glm(Status ~ (Hub13_B_Cell + Hub2_T_Cell + Hub4_T_Cell+
                               Hub10_T_Cell + Hub11_T_Cell + Hub13_T_Cell +
                               Hub16_T_Cell + Hub18_T_Cell + CYTO1 + CYTO2 +
                               CYTO3), data = df1,
                family = "binomial")
summary(logistic_rf)
step_rf <- step(logistic_rf)
summary(step_rf)
pchisq(step_rf$deviance, step_rf$df.residual, lower.tail = FALSE)
#This stepwise model is significant#
#We found 5 significant variables after using stepwise: Hub2_T_Cell, Hub10_T_Cell, Hub11_T_Cell, Hub18_T_Cell, CYTO1, CYTO2, CYTO3. These are the variables which influence outcome of MRSA individually the most.#


##together##
logistic_rf_tog <- glm(Status ~ ((Hub2_T_Cell + Hub10_T_Cell + Hub11_T_Cell + 
                                    Hub18_T_Cell)*(CYTO1 + CYTO2 + CYTO3)),
                       data = df1, family = "binomial")
summary(logistic_rf_tog)
anova(logistic_rf, logistic_rf_tog, test = "Chi")
##The interaction terms are not significant. These hub genes and cytokines does not influence outcome of MRSA together. The result matches with lasso method.##



```

c. Are any of these hub genes and cytokines significantly different between two groups?
(two sample t-test? normality?)
## Boxplots comparing each of the variable distributions between resolving (0) and persiting (1)

```{r}
library(tidyverse)

df_main <- MRSA_status %>% 
  left_join(B_cell, by = c("SID" = "SID")) %>% 
  left_join(T_cell, by = c("SID" = "SID"), suffix = c("_B_Cell", "_T_Cell")) %>% 
  left_join(Cytokine, by = c("SID" = "SID"))

df_main$Status <- as.factor(df_main$Status)

#B_Cell
df_Bcell <- df_main %>% 
  dplyr::select(SID, Status, ends_with("B_Cell")) %>%
  pivot_longer(!c(SID, Status), names_to = "B_cell", values_to = "Measurement")

ggplot(df_Bcell, aes(x = "Status", y = Measurement, fill = Status)) + 
    geom_boxplot() +
    facet_wrap(~B_cell, scale = "free")

#T_cell
df_Tcell <- df_main %>% 
  dplyr::select(SID, Status, ends_with("T_Cell")) %>%
  pivot_longer(!c(SID, Status), names_to = "T_cell", values_to = "Measurement")

ggplot(df_Tcell, aes(x = "Status", y = Measurement, fill = Status)) + 
    geom_boxplot() +
    facet_wrap(~T_cell, scale = "free")

#Cytokine
df_Cyto <- df_main %>% 
  dplyr::select(SID, Status, starts_with("CYTO")) %>%
  pivot_longer(!c(SID, Status), names_to = "Cytokine", values_to = "Measurement")

ggplot(df_Cyto, aes(x = "Status", y = Measurement, 
                             fill = Status)) + 
    geom_boxplot() +
    facet_wrap(~Cytokine, scale = "free")
```

## Transform by adding 1 then logging (to meet upcoming tests' normality assumption)

```{r}
df_main_trans <- df_main 
df_main_trans[, 3:45] <- log(df_main_trans[, 3:45] + 1)
```

## Two sample t-tests to compare geometric means of each variable between resolving (0) and persiting (1)

```{r}
#B_CELL 1

#test for equal variances: VARIANCES EQUAL
var.test(Hub1_B_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub1_B_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub1_B_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#B_CELL 2

#test for equal variances: VARIANCES EQUAL
var.test(Hub2_B_Cell ~ Status, data = df_main_trans)

#two sample t-test: SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub2_B_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub2_B_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#B_CELL 3

#test for equal variances: VARIANCES EQUAL
var.test(Hub3_B_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub3_B_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub3_B_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#B_CELL 4

#test for equal variances: VARIANCES EQUAL
var.test(Hub4_B_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub4_B_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub4_B_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#B_CELL 5

#test for equal variances: VARIANCES EQUAL
var.test(Hub5_B_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub5_B_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub5_B_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#B_CELL 6

#test for equal variances: VARIANCES EQUAL
var.test(Hub6_B_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub6_B_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub6_B_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#B_CELL 7

#test for equal variances: VARIANCES EQUAL
var.test(Hub7_B_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub7_B_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub7_B_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#B_CELL 8

#test for equal variances: VARIANCES EQUAL
var.test(Hub8_B_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub8_B_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub8_B_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#B_CELL 9

#test for equal variances: VARIANCES EQUAL
var.test(Hub9_B_Cell ~ Status, data = df_main_trans)

#two sample t-test: SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub9_B_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub9_B_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#B_CELL 10

#test for equal variances: VARIANCES EQUAL
var.test(Hub10_B_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub10_B_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub10_B_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#B_CELL 11

#test for equal variances: VARIANCES EQUAL
var.test(Hub11_B_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub11_B_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub11_B_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#B_CELL 12

#test for equal variances: VARIANCES EQUAL
var.test(Hub12_B_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub12_B_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub12_B_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#B_CELL 13

#test for equal variances: VARIANCES EQUAL
var.test(Hub13_B_Cell ~ Status, data = df_main_trans)

#two sample t-test: SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub13_B_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub13_B_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#B_CELL 14

#test for equal variances: VARIANCES EQUAL
var.test(Hub14_B_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub14_B_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub14_B_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#B_CELL 15

#test for equal variances: VARIANCES EQUAL
var.test(Hub15_B_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub15_B_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub15_B_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#B_CELL 16

#test for equal variances: VARIANCES EQUAL
var.test(Hub16_B_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub16_B_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub16_B_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#B_CELL 17

#test for equal variances: VARIANCES EQUAL
var.test(Hub17_B_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub17_B_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub17_B_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#B_CELL 18

#test for equal variances: VARIANCES EQUAL
var.test(Hub18_B_Cell ~ Status, data = df_main_trans)

#two sample t-test: SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub18_B_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub18_B_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#B_CELL 19

#test for equal variances: VARIANCES EQUAL
var.test(Hub19_B_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub19_B_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub19_B_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#B_CELL 20

#test for equal variances: VARIANCES EQUAL
var.test(Hub20_B_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub20_B_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub20_B_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#T_CELL 1

#test for equal variances: VARIANCES EQUAL
var.test(Hub1_T_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub1_T_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub1_T_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#T_CELL 2

#test for equal variances: VARIANCES EQUAL
var.test(Hub2_T_Cell ~ Status, data = df_main_trans)

#two sample t-test: SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub2_T_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub2_T_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#T_CELL 3

#test for equal variances: VARIANCES EQUAL
var.test(Hub3_T_Cell ~ Status, data = df_main_trans)

#two sample t-test: SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub3_T_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub3_T_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#T_CELL 4

#test for equal variances: VARIANCES EQUAL
var.test(Hub4_T_Cell ~ Status, data = df_main_trans)

#two sample t-test: SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub4_T_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub4_T_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#T_CELL 5

#test for equal variances: VARIANCES EQUAL
var.test(Hub5_T_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub5_T_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub5_T_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#T_CELL 6

#test for equal variances: VARIANCES EQUAL
var.test(Hub6_T_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub6_T_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub6_T_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#T_CELL 7

#test for equal variances: VARIANCES EQUAL
var.test(Hub7_T_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub7_T_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub7_T_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#T_CELL 8

#test for equal variances: VARIANCES EQUAL
var.test(Hub8_T_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub8_T_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub8_T_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#T_CELL 9

#test for equal variances: VARIANCES EQUAL
var.test(Hub9_T_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub9_T_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub9_T_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#T_CELL 10

#test for equal variances: VARIANCES EQUAL
var.test(Hub10_T_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub10_T_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub10_T_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#T_CELL 11

#test for equal variances: VARIANCES UNEQUAL
var.test(Hub11_T_Cell ~ Status, data = df_main_trans)

#two sample t-test: SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub11_T_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub11_T_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = FALSE)
```

```{r}
#T_CELL 12

#test for equal variances: VARIANCES EQUAL
var.test(Hub12_T_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub12_T_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub12_T_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#T_CELL 13

#test for equal variances: VARIANCES UNEQUAL
var.test(Hub13_T_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub13_T_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub13_T_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = FALSE)
```

```{r}
#T_CELL 14

#test for equal variances: VARIANCES UNEQUAL
var.test(Hub14_T_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub14_T_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub14_T_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = FALSE)
```

```{r}
#T_CELL 15

#test for equal variances: VARIANCES EQUAL
var.test(Hub15_T_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub15_T_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub15_T_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#T_CELL 16

#test for equal variances: VARIANCES UNEQUAL
var.test(Hub16_T_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub16_T_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub16_T_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = FALSE)
```

```{r}
#T_CELL 17

#test for equal variances: VARIANCES EQUAL
var.test(Hub17_T_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub17_T_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub17_T_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#T_CELL 18

#test for equal variances: VARIANCES EQUAL
var.test(Hub18_T_Cell ~ Status, data = df_main_trans)

#two sample t-test: SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub18_T_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub18_T_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#T_CELL 19

#test for equal variances: VARIANCES EQUAL
var.test(Hub19_T_Cell ~ Status, data = df_main_trans)

#two sample t-test: SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub19_T_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub19_T_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#T_CELL 20

#test for equal variances: VARIANCES EQUAL
var.test(Hub20_T_Cell ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$Hub20_T_Cell[df_main_trans$Status == 0], 
       df_main_trans$Hub20_T_Cell[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#CYTO 1

#test for equal variances: VARIANCES UNEQUAL
var.test(CYTO1 ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$CYTO1[df_main_trans$Status == 0], 
       df_main_trans$CYTO1[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = FALSE)
```

```{r}
#CYTO 2

#test for equal variances: VARIANCES EQUAL
var.test(CYTO2 ~ Status, data = df_main_trans)

#two sample t-test: SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$CYTO2[df_main_trans$Status == 0], 
       df_main_trans$CYTO2[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

```{r}
#CYTO 3

#test for equal variances: VARIANCES EQUAL
var.test(CYTO3 ~ Status, data = df_main_trans)

#two sample t-test: NOT SIGNIFICANTLY DIFFERENT
t.test(df_main_trans$CYTO3[df_main_trans$Status == 0], 
       df_main_trans$CYTO3[df_main_trans$Status == 1], 
       alternative = "two.sided", var.equal = TRUE)
```

Significant differences were found between the geometric means for resolving and persisting for:
-bhub2
-bhub9
-bhub13
-bhub18
-thub2
-thub3
-thub4
-thub11
-thub18
-thub19
-cyto2
-cyto3

2. What hub genes and cytokines in the blood sample specifically correlate with: 
a. Resolving outcome of MRSA infection (ARMB)
b. Persistent outcome of MRSA infection (APMB)

```{r}
library(corrplot)
library(dplyr)

df <- subset(df, select = -c(SID) )

df_Status_0_Tcell <- df %>% filter(Status == 0) %>% dplyr::select(ends_with("T_cell"))
df_Status_1_Tcell <- df %>% filter(Status == 1) %>% dplyr::select(ends_with("T_cell"))
df_Status_0_Bcell <- df %>% filter(Status == 0) %>% dplyr::select(ends_with("B_cell"))
df_Status_1_Bcell <- df %>% filter(Status == 1) %>% dplyr::select(ends_with("B_cell"))
df_Status_0_Cyto <- df %>% filter(Status == 0) %>% dplyr::select(starts_with("CYTO"))
df_Status_1_Cyto <- df %>% filter(Status == 1) %>% dplyr::select(starts_with("CYTO"))

cor_matrix_Tcell_0 <- cor(df_Status_0_Tcell)
cor_matrix_Tcell_1 <- cor(df_Status_1_Tcell)
cor_matrix_Bcell_0 <- cor(df_Status_0_Bcell)
cor_matrix_Bcell_1 <- cor(df_Status_1_Bcell)
cor_matrix_Cyto_0 <- cor(df_Status_0_Cyto)
cor_matrix_Cyto_1 <- cor(df_Status_1_Cyto)

'
corrplot(cor_matrix_Tcell_0, method="color", type="upper", order="hclust", tl.col="black")
corrplot(cor_matrix_Tcell_1, method="color", type="upper", order="hclust", tl.col="black")
corrplot(cor_matrix_Bcell_0, method="color", type="upper", order="hclust", tl.col="black")
corrplot(cor_matrix_Bcell_1, method="color", type="upper", order="hclust", tl.col="black")
corrplot(cor_matrix_Cyto_0, method="color", type="upper", order="hclust", tl.col="black")
corrplot(cor_matrix_Cyto_1, method="color", type="upper", order="hclust", tl.col="black")
'

cor.mtest <- function(mat, ...) {
    mat <- as.matrix(mat)
    n <- ncol(mat)
    p.mat<- matrix(NA, n, n)
    diag(p.mat) <- 0
    for (i in 1:(n - 1)) {
        for (j in (i + 1):n) {
            tmp <- cor.test(mat[, i], mat[, j], ...)
            p.mat[i, j] <- p.mat[j, i] <- tmp$p.value
        }
    }
  colnames(p.mat) <- rownames(p.mat) <- colnames(mat)
  p.mat
}

p.mat_T0 <- cor.mtest(df_Status_0_Tcell)
p.mat_T1 <- cor.mtest(df_Status_1_Tcell)
p.mat_B0 <- cor.mtest(df_Status_0_Bcell)
p.mat_B1 <- cor.mtest(df_Status_1_Bcell)
p.mat_Cyto0 <- cor.mtest(df_Status_0_Cyto)
p.mat_Cyto1 <- cor.mtest(df_Status_1_Cyto)

corrplot(cor_matrix_Tcell_0, method="color", type="upper", order="hclust", tl.col="black", p.mat = p.mat_T0, sig.level = 0.05, insig = "blank", title="Tcell_0", mar=c(0,0,1,0))

corrplot(cor_matrix_Tcell_1, method="color", type="upper", order="hclust", tl.col="black", p.mat = p.mat_T1, sig.level = 0.05, insig = "blank", title="Tcell_1", mar=c(0,0,1,0))

corrplot(cor_matrix_Bcell_0, method="color", type="upper", order="hclust", tl.col="black", p.mat = p.mat_B0, sig.level = 0.05, insig = "blank", title="Bcell_0", mar=c(0,0,1,0))

corrplot(cor_matrix_Bcell_1, method="color", type="upper", order="hclust", tl.col="black", p.mat = p.mat_B1, sig.level = 0.05, insig = "blank", title="Bcell_1", mar=c(0,0,1,0))

corrplot(cor_matrix_Cyto_0, method="color", type="upper", order="hclust", tl.col="black", p.mat = p.mat_Cyto0, sig.level = 0.05, insig = "blank", title="Cyto_0", mar=c(0,0,1,0))

corrplot(cor_matrix_Cyto_1, method="color", type="upper", order="hclust", tl.col="black", p.mat = p.mat_Cyto1, sig.level = 0.05, insig = "blank", title="Cyto_1", mar=c(0,0,1,0))
```