---
title: "Biostat 402A Assn2"
author: "Yenlin Lai"
date: "4/14/2022"
output: pdf_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```


```{r}
library(haven)
bpd <- read_dta("D:/UCLA Biostat/2022 Spring/Biostat 402A/Assignment2/bpd.dta")

library(dplyr)
bpd <- bpd %>% 
  as_tibble() %>% 
  print(width = Inf)
```

3. Perform a descriptive analysis including the following: 
a. A demographic table describing the study population.
```{r}
library(table1)
bpd$sex <- 
  factor(bpd$sex, levels=c(1,0),
         labels=c("Male", 
                  "Female"))
bpd$agar <- 
  factor(bpd$agar)

label(bpd$sex) <- "Sex"
label(bpd$bxy) <- "Birth year"
label(bpd$agar) <- "APGAR at 1 minute"
label(bpd$gage) <- "Gestation age"
label(bpd$bwt) <- "Birthweight"
label(bpd$rage) <- "Respiratory symptoms age"
label(bpd$vage) <- "Ventilatory assistance age"

units(bpd$gage) <- "weeks*10"
units(bpd$bwt) <- "gm"
units(bpd$rage) <- "hours*10"
units(bpd$vage) <- "hours"

table1(~ sex + bxy + agar + gage + bwt + rage + vage, data = bpd,
        render.continuous = c("Median" = "Median", "IQR" = "IQR"))
```

b. Plots (histograms or box plots) and summary statistics for the distributions of: 
i. hours of exposure to low concentrations of elevated oxygen 
ii. hours of exposure to medium concentrations of elevated oxygen 
iii. hours of exposure to high concentrations of elevated oxygen 
iv. hours of duration of endotracheal intubation 
v. hours of duration of assisted ventilation 
```{r}
library(ggplot2)
library(reshape2)
bpd_b <- bpd %>%
  select(lo2, mo2, ho2, idur, vdur) %>% 
  melt()

bpd_b %>% 
  ggplot(aes(x = variable, y = value, fill = variable)) +
  geom_boxplot() +
  coord_cartesian(ylim = quantile(bpd_b$value, c(0,0.99), na.rm = TRUE)) +
  labs(y = "Hours", x = "") +
  scale_fill_brewer(palette = "Blues") + 
  theme(legend.position = "none") + 
  theme_light()

label(bpd$lo2) <- "Exposure to low concentrations of elevated oxygen"
label(bpd$mo2) <- "Exposure to medium concentrations of elevated oxygen "
label(bpd$ho2) <- "Exposure to high concentrations of elevated oxygen"
label(bpd$idur) <- "Duration of endotracheal intubation"
label(bpd$vdur) <- "Duration of assisted ventilation"

units(bpd$lo2) <- "hours"
units(bpd$mo2) <- "hours"
units(bpd$ho2) <- "hours"
units(bpd$idur) <- "hours"
units(bpd$vdur) <- "hours"

table1(~ lo2 + mo2 + ho2 + idur + vdur, data = bpd)
```
c. Counts (or bar graphs) of RDS and radiographic BPD scores.
```{r}
ggplot(bpd, aes(x = factor(rds))) +
  geom_bar() + 
  geom_text(aes(label = ..count..), stat = "count", vjust = -0.2) +
  labs(x = "Severity of RDS", title = "Counts for RDS Severity") + 
  scale_fill_brewer(palette = "Blues") +
  theme_light()

bpd <- bpd %>% 
  mutate(bpd = bpd/10)
ggplot(bpd, aes(x = factor(bpd))) +
  geom_bar() + 
  geom_text(aes(label = ..count..), stat = "count", vjust = -0.2) +
  labs(x = "Radiographic BPD score", title = "Counts for Radiographic BPD") +
  theme_light()

bpd <- bpd %>% 
  mutate(haverds = ifelse(rds >= 1, "Yes", "No"),
         severerds = ifelse(rds == 5, "Yes", "No"),
         haveBPD = ifelse(bpd >=3, "Yes", "No"))
bpd %>% 
  ggplot(aes(x = haverds)) +
  geom_bar() + 
  geom_text(aes(label = ..count..), stat = "count", vjust = -0.2) +
  labs(x = "Infants had RDS", title = "Bar chart for Infants with RDS") +
  theme_light()

bpd %>% 
  ggplot(aes(x = severerds)) +
  geom_bar() + 
  geom_text(aes(label = ..count..), stat = "count", vjust = -0.2) +
  labs(x = "Infants had Severe RDS", 
       title = "Bar chart for Infants with Severe RDS") +
  theme_light()

bpd %>% 
  ggplot(aes(x = haveBPD)) +
  geom_bar() + 
  geom_text(aes(label = ..count..), stat = "count", vjust = -0.2) +
  labs(x = "Infants had BPD", title = "Bar chart for Infants with BPD") +
  theme_light()
  

```


d. Missing data and outliers. Identify variables with large numbers of missing data values (say, >20% missing). List variables with extreme outliers.

```{r}
library(naniar)
bpd <- 
  bpd %>% 
  replace_with_na(replace = list(brmu = 0, puin = 0, hyme = 0, alin = 0,
                                 emph = 0, lbpd = 0, hema = 0))
bpd %>% 
  sapply(function(x) sum(is.na(x)/length(x)))

bpd %>%
  select(-c(haverds, severerds, haveBPD)) %>% 
  sapply(function(x) sort(boxplot(x, plot = FALSE)$out))
```